---
title: "Using `WAACHShelp::icd_morb_flag` in practice"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using `WAACHShelp::icd_morb_flag` in practice}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, echo = F}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Setup

```{r setup, echo = T, message=F, warning=FALSE}
# Load the packages
library(WAACHShelp)
```

# Data requirements

`icd_morb_flag` has two dataframe inputs: `data`, and `dobmap`.

+ `data`
  + A hospital morbidity dataset.
  + We require of this dataset is:
    + Some sort of ID value (to differentiate visits within an individual)
    + A date variable (only if flagging whether a visit occurred below a certain age)
    + The set of flagging variable(s) to search for ICD codes across. It does not matter what these flagging variables are called.
+ `dobmap`
  + A DOBmap file.
  + In reality, we only need two columns: some sort of ID value, and a dob value.

The ID values must have the same name across the `data` and `dobmap` files. 

By default, we assume the `rootnum` variable exists in both data sets. It does not matter what this variable is called in reality---the `joining_var` argument can be specified to any string.

# Flagging

Flagging can be handled by automatic or manual specification, via `flag_category`
